---
title: "Ocean Sciences Figures"
author: "Mukta Kelkar"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)

#tidyverse packages
library(here)
library(janitor)
library(dplyr)
library(tidyverse)

#spatial data
library(sf)

#datavis
library(RColorBrewer)
library(showtext)
library(pals)
library(rcartocolor)
```

```{r read in & clean data}
#read in lit review data
data <- read.csv(here("data10.csv")) %>% 
  clean_names() %>% 
  select(-mpa_effect_size) #not currently considering this variable

#read in MEOW TNC Shapefile
meow <- read_sf(here("MEOW-TNC", "meow_ecos.shp")) %>% 
  clean_names()

#add google fonts to working memory for figures
font_add_google(name = "Work Sans",   # Name of the font on the Google Fonts site
                family = "worksans")

#standard font size for all graphs
global_size = 14
```


# Set of pie charts for Figure 1
Make a pie chart for each of the following parameters:
- habitat_type
- climate_effect
- mpa_effect
- marine_ecoregion


```{r figure 1}
#figure 1a
habitat_counts <- data %>% 
  count(habitat_type) #make dataset with counts data

fig1a <- ggplot(data = habitat_counts, aes (x = "", y = n, fill = habitat_type)) +
  geom_bar(stat = "identity", width = 1) +   #set up bar chart
  coord_polar("y") +                                          #put bar chart in polar coordinates to shape it into a pie chart
  theme_void() +                                              #get rid of background
  theme(legend.position = "right",
        legend.text = element_text(family = "Arial", size = 14, color = "#001b40"),
        legend.title = element_blank()) + 
  labs(
    fill = "Habitat Type"
  ) +
  scale_fill_carto_d(palette = "Safe",
                     labels = c("Coral Reef", "Deep Reef", "NA", "Rocky Intertidal", "Rocky Reef", "Seagrass", "Shallow Reef", "Subtropical Reef", "Temperate Reef", "Temperate Transition Zone" )) +
  #scale_fill_manual(values = c("mediumpurple3", "chocolate3","#CDCDC1", "blue3","forestgreen","orangered1", "orange", "#ee6aa7", "maroon", "gray21")) +
  guides(color = none)
  

#figure 1b
mpa_effect_counts <- data %>% 
  count(mpa_effect) #make dataset with counts data

fig1b <- ggplot(data = mpa_effect_counts, aes (x = "", y = n, fill = mpa_effect)) +
  geom_bar(stat = "identity", width = 1) +   #set up bar chart
  coord_polar("y") +                                          #put bar chart in polar coordinates to shape it into a pie chart
  theme_void() +                                              #get rid of background
  theme(legend.position = "right",
        legend.text = element_text(family = "Arial", size = 14, color = "#001b40"),
        legend.title = element_blank()) + 
  labs(
    fill = "MPA Effect Type"
  ) +
  scale_fill_brewer(palette = "PuBu",
                    labels = c("Ecosystem (+)", "Population", "Ecosystem (-)", "None")) +
  #scale_fill_manual(values = c("#CD2626","forestgreen","#1874CD", "#EE9A49", "#8B8682"),
                    #labels = c("Direct", "Support", "Offset", "Negative Effect", "No Effect")) +
  guides(color = none)

#figure 1c
ecoregion_counts <- data %>% 
  count(marine_ecoregion) #make dataset with counts data

fig1c <- ggplot(data = ecoregion_counts, aes (x = "", y = n, fill = marine_ecoregion)) +
  geom_bar(stat = "identity", width = 1, color = "white") +   #set up bar chart
  coord_polar("y") +                                          #put bar chart in polar coordinates to shape it into a pie chart
  theme_void() +                                              #get rid of background
  theme(legend.position = "right",
        legend.text = element_text(family = "Arial", size = 12, color = "#001b40")) + 
  labs(
    fill = "Marine Ecoregion"
  )

#figure 1d
climate_counts_prelim <- data %>% 
  count(climate_effect) #preliminary counts of climate effect (without separating rows with multiple categories)

climate_counts <- climate_counts_prelim %>%  #read in preliminary counts
  separate_rows(climate_effect, sep = ";") %>%  #separate rows along ";"
  mutate(n = n*1) %>%   #repeat count values for each separated row
  group_by(climate_effect) %>%   #group by climate effect name to combine duplicates
  summarize(total_count = sum(n)) #find the sum of each category identified in `group_by()`

fig1d <- ggplot(data = climate_counts, aes (x = "", y = total_count, fill = climate_effect)) +
  geom_bar(stat = "identity", width = 1, color = "white") +   #set up bar chart
  coord_polar("y") +                                          #put bar chart in polar coordinates to shape it into a pie chart
  theme_void() +                                              #get rid of background
  theme(legend.position = "right") + 
  labs(
    fill = "Climate Effect"
  )
  

#print figures
fig1a
fig1b
fig1c
fig1d

#save figures
ggsave("fig1a.png", plot = fig1a, width = 9, height = 5, bg = "transparent")
ggsave("fig1b.png", plot = fig1b, width = 9, height = 5, bg = "transparent") #save fig1b with a transparent background

```

```{r histograms and barplot versions of the same data}
#figure 1a_bar
fig1a_bar <- ggplot(data = habitat_counts, aes (x = reorder(habitat_type, -n), y = n)) +
  geom_bar(stat = "identity", width = .75) +   #set up bar chart
  theme_classic(base_size = global_size) +
  labs(
    x = "Habitat Type",
    y = "Number of Occurences"
  ) +
  #scale_x_discrete(labels = c("Coral Reef", "Deep Reef", "NA", "Rocky Intertidal", "Rocky Reef", #"Seagrass", "Shallow Reef", "Subtropical Reef", "Temperate Reef", "Temperate Transition Zone")) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
  

#figure 1b
fig1b_bar <- ggplot(data = mpa_effect_counts, aes (x = reorder(mpa_effect, -n), y = n)) +
  geom_bar(stat = "identity", width = .75) +   #set up bar chart
  theme_classic(base_size = global_size) +
  labs(
    x = "MPA Effect Type",
    y = "Number of Occurences"
  ) +
  scale_x_discrete(labels = c("None", "Ecosystem (+)", "Population", "Ecosystem (-)")) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
  

#figure 1c
fig1c_bar <- ggplot(data = ecoregion_counts, aes (x = reorder(marine_ecoregion, -n), y = n)) +
  geom_bar(stat = "identity", width = .75, color = "white") +   #set up bar chart
  theme_classic(base_size = global_size) +
  labs(
    x = "Ecoregion",
    y = "Number of Occurences"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )


#figure 1d
fig1d_bar <- ggplot(data = climate_counts, aes (x = reorder(climate_effect, -total_count), y = total_count)) +
  geom_bar(stat = "identity", width = .75) +   #set up bar chart
  theme_classic(base_size = global_size) +
  labs(
    x = "Ecological Effect of Climate Change",
    y = "Number of Occurences"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
  

#print figures
fig1a_bar
fig1b_bar
fig1c_bar
fig1d_bar
```



# Maps for Figure 2
- https://stackoverflow.com/questions/10368180/plotting-pie-graphs-on-map-in-ggplot

Next Steps:
- make the ecoregions grayscale
- turn the bubbles into pie charts
-- MPA effects pie charts on one map
-- ecosystem types pie chart on the other map


```{r geography, message = FALSE, warning = FALSE}
#bar plot of country locations
ggplot(data = ecoregion_counts, aes(x = marine_ecoregion, y = n)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=.4))

#make spatial polygon for marine realms
realms <- meow %>%
  group_by(realm) %>%
  summarize(geometry = st_combine(geometry)) %>% 
  mutate(realm = recode(realm,
                        "Arctic" = "arctic",
                        "Central Indo-Pacific" = "central_indo_pacific",
                        "Eastern Indo-Pacific" = "eastern_indo_pacific",
                        "Southern Ocean" = "southern_ocean",
                        "Temperate Australasia" = "temperate_autralasia",
                        "Temperate Northern Atlantic" = "temperate_northern_atlantic",
                        "Temperate Northern Pacific" = "temperate_northern_pacific",
                        "Temperate South America" = "temperate_south_america",
                        "Temperate Southern Africa" = "temperate_southern_africa",
                        "Tropical Atlantic" = "tropical_atlantic",
                        "Tropical Eastern Pacific" = "tropical_eastern_pacific",
                        "Western Indo-Pacific" = "western_indo_pacific")) %>% 
  rename(marine_ecoregion = realm)

#create manual coordinates for temperate_northern_pacific centroid
new_coordinates <- st_point(c(-125, 38))

#find centroids of each ecoregion
symbol_pos <- st_centroid(realms, of_largest_polygon = TRUE) %>% 
  assign_in(list(2,7), new_coordinates) #manually replace temperate_northern_pacific centroid

#df <- symbol_pos %>% 
 # filter(marine_ecoregion == "temperate_northern_pacific") %>% 
  #mutate(geometry = new_coordinates)

#join centroids of each ecoregion to counts data, so that the counts data can be placed at the centroid on the map
count_n <- left_join(symbol_pos, ecoregion_counts, by = "marine_ecoregion")

#change NAs to 0s
count_n$n[is.na(count_n$n)] <- 0

#make a plot
#refining steps: need to add labels to the bubbles, need to figure out ideal color for bubbles
marine_realms <- ggplot() +
  geom_sf(data = realms, aes(fill = marine_ecoregion, color = NA)) +
  scale_fill_manual(values = viridis::viridis(length(unique(realms$marine_ecoregion)))) + #specifies colors from virdis package manually. this is a workaround to avoid getting the error " Discrete value supplied to continuous scale" when using `scale_fill_viridis`. `scale_fill_gray` worked because its more flexible
  #scale_color_manual(values = viridis::viridis(length(unique(realms$marine_ecoregion)))) +
  labs(
    fill = "",
    caption = "This map does not include 4 global studies and 5 studies in the northern hemisphere."
  ) +
  geom_sf(data = count_n, 
          pch = 21, 
          aes(size = n,
              color = "red"),
          fill = "red",
          show.legend = FALSE) +
  scale_size(range = c(0, 9)) +      #adjust size of bubbles on map
  geom_text(data = count_n, #add a geom for labels of the bubble sizes
            aes(x = st_coordinates(geometry)[,1], y = st_coordinates(geometry)[, 2],   #add x and y values to geom_text
              nudge_x = 0.5, nudge_y = 0.5, label = n)) +    #add data labels to each bubble
  labs(
    x = "",
    y = ""
  ) +
  theme_classic() +
  guides(fill = guide_legend(key_glyph = "point", override.aes = list(color = "white")),  #set legend key as a point, and then change the legend key color to white
         color = FALSE) + #don't show color legend
  ggtitle("Number of case studies conducted in each marine realm") +
  theme(plot.title = element_text(hjust = 0.5))   #move title to center

marine_realms
```

## MPA Effects by Region

## Habitat Types by Region

# Mechanism graph for Figure 3

```{r figure 3}
mechanisms <- data %>% 
  select(mechanism_key_action_leads_to_consequence) %>% 
  count(mechanism_key_action_leads_to_consequence)

```

Table of MPA Effects and Corresponding Mechanisms

```{r}
#effect a
a_mech <- data %>% 
  filter(mpa_effect == "effect_a") %>% 
  select(mpa_effect, mechanism_key_action_leads_to_consequence) %>% 
  count(mechanism_key_action_leads_to_consequence)

#effect b
b_mech <- data %>% 
  filter(mpa_effect == "effect_b") %>% 
  select(mpa_effect, mechanism_key_action_leads_to_consequence) %>% 
  count(mechanism_key_action_leads_to_consequence)

c_mech <- data %>% 
  filter(mpa_effect == "effect_c") %>% 
  select(mpa_effect, mechanism_key_action_leads_to_consequence) %>% 
  count(mechanism_key_action_leads_to_consequence)

no_effect_mech <- data %>% 
  filter(mpa_effect == "no_effect") %>% 
  select(mpa_effect, mechanism_key_action_leads_to_consequence) %>% 
  count(mechanism_key_action_leads_to_consequence)

negative_mech <- data %>% 
  filter(mpa_effect == "negative_effect") %>% 
  select(mpa_effect, mechanism_key_action_leads_to_consequence) %>% 
  count(mechanism_key_action_leads_to_consequence)


```



```{r eval = FALSE}
#save list of mpa effects
effect_list <- unique(data$mpa_effect)

# create empty dataframe
mech_effects <- data.frame(matrix(ncol = 3, nrow = nrow(mechanisms)))

#provide column names
colnames(mech_effects) <- c("mpa_effect", "mechanism", "count")

for(i in 1:length(effect_list)) {  #for every element starting with the first element
  
  #filter by mpa effect and save to holding vector `effect`
  effect <- data %>% filter(mpa_effect == effect_list[i]) 
  
  #make list of all mechanisms and save to holding vector `mech`
  mech <- unique(effect$habitat_type)   
  
  #delete nas, save to holding vector `habs_no_na`
  habs_no_na <- toString(as.list(habs[!habs == "na"])) #save unique values as a list, and then put the entire list in one cell of the holding vector
  
  #make a list of all marine realms holding vector `re`
  re <- toString(as.list(unique(effect$marine_ecoregion))) #save unique values as a list, and then put the entire list in one cell of the holding vector
  
  #create a holding vector to append as a row to the dataframe
  vec <- c((effect_list[i]),(habs_no_na), (re)) # create vector of mpa effect, habitat, and marine ecoregion
  
  #append rows to dataframe
  mpa_habitats[i,] <- vec
}


#drop rows where mpa_effect cell is blank (delete this line later)
mpa_eff_habitats <- mpa_habitats %>% 
  filter(!(mpa_effect == ""))

#make table
effects_table <- mpa_eff_habitats %>% 
  kbl() %>% 
  kable_styling("hover", full_width = F)

#print table
effects_table
```


# Figure 4
Climate Effects Bar Chart

```{r}
eco_effects <- data %>% 
  select(mpa_effect, climate_effect) %>% 
  separate_rows(climate_effect, sep = ";") %>%   #separate rows along ";"
  group_by(mpa_effect, climate_effect) %>% 
  count(mpa_effect, climate_effect, name = "count")

climateeffects_barchart <- ggplot(data = eco_effects, aes(fill = climate_effect, y = count, x = mpa_effect)) +
  geom_col() +
  scale_x_discrete(labels = c("Ecosystem (+)", "Population", "Ecosystem (-)", "None")) +
  theme_minimal() +
  labs(
    y = "",
    x = "",
    fill = "Climate Effects"
  ) +
  scale_fill_manual(values = unname(kelly()),
                    labels = c("Calcification", "Disease", "Habitat Availability", "Nutrition", "Phenology", "Range Shift", "Reproduction")) +
  #scale_fill_manual(values = c("#FFF5EE", "#FFA500", "#68228B", "royalblue3", "#228B22", "#CD1076")) +
  theme(
    axis.text.x = element_text(size = 14, family = "Arial", color = "#001b40", angle = 30, vjust = .75),
    axis.text.y = element_text(size = 14, family = "Arial", color = "#001b40"),
    legend.text = element_text(size = 14, family = "Arial", color = "#001b40"),
    legend.title = element_text(size = 14, family = "Arial", color = "#001b40"),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank())

climateeffects_barchart

ggsave("climateeffects_barchart.png", plot = climateeffects_barchart, width = 10, height = 5, bg = "transparent")

```


# Heat Map for Figure 5
```{r}
heatmap1 <- ggplot(data = eco_effects, aes(x = climate_effect, y = mpa_effect, fill = count)) +
  geom_tile() +
  labs(
    x = "Climate Effect",
    y = "MPA Effect"
  ) +
  scale_y_discrete(labels = c("Ecosystem (+)", "Population", "Ecosystem (-)", "None")) +
  scale_fill_gradient2(low = "grey",
                      high = "black") +
  theme_classic(base_size = global_size) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

heatmap1


ggsave("heatmap1.png", plot = heatmap1, width = 10, height = 5, bg = "transparent")
```


# Heat Map for Figure 6
```{r}
#look at climate and mpa effects by habitat
habitat_climate <- data %>% 
  select(habitat_type, climate_effect) %>% 
  separate_rows(climate_effect, sep = ";") %>%   #separate rows along ";"
  group_by(climate_effect, habitat_type) %>% 
  count(climate_effect, habitat_type, name = "count")

heatmap2 <- ggplot(data = habitat_climate, aes(x = climate_effect, y = habitat_type, fill = count)) +
  geom_tile() +
  labs(
    x = "Climate Effect",
    y = "Habitat Type"
  ) +
  scale_fill_gradient2(low = "grey",
                      high = "black") +
  theme_classic(base_size = global_size) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

heatmap2

ggsave("heatmap2.png", plot = heatmap2, width = 10, height = 5, bg = "transparent")

```

# Heat Map for Figure 7
```{r}
mpa_habitat <- data %>% 
  select(mpa_effect, habitat_type) %>%
  group_by(mpa_effect, habitat_type) %>% 
  count(mpa_effect, habitat_type, name = "count")

heatmap3 <- ggplot(data = mpa_habitat, aes(x = mpa_effect, y = habitat_type, fill = count)) +
  geom_tile() +
   labs(
    x = "MPA Effect",
    y = "Habitat Type"
  ) +
  scale_x_discrete(labels = c("Ecosystem (+)", "Population", "Ecosystem (-)", "None")) +
  scale_fill_gradient2(low = "grey",
                      high = "black") +
  theme_classic(base_size = global_size)

heatmap3
  
ggsave("heatmap3.png", plot = heatmap3, width = 10, height = 5, bg = "transparent")
```


#Archived Notes

```{r, eval = FALSE}
# Sample data
sample <- tribble(
  ~item, ~n,
  "a", 1,
  "a;b", 2,
  "b", 1,
  "c", 1,
  "a;c", 3
)

# Separate the items by semicolon and replicate the counts
result <- sample %>%
  separate_rows(item, sep = ";") %>%
  mutate(n = n * 1)  # Replicate counts

# Display the result
print(result)


# Example data
sample <- data.frame(
  cat1 = c("a", "a", "b", "c", "c"),
  cat2 = c("cat", "dog", "rabbit", "cat", "cat")
)

# Example plot with equal width bars
p <- ggplot(data = sample, aes(x = cat1, fill = cat2)) +
  geom_bar(position = "dodge", width = 0.5) + # Adjust width as needed
  labs(title = "Sample Plot", x = "Category 1", y = "Count") +
  scale_fill_manual(values = c("cat" = "blue", "dog" = "green", "rabbit" = "red"),
                    labels = c("Cat Label", "Dog Label", "Rabbit Label")) +
  theme(legend.title = element_blank(),
        legend.key = element_blank())  # Remove legend border

# Display plot
print(p)

```


