---
title: "Explore Data for Climate Resilient MPAs Lit Review"
author: "Mukta Kelkar"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
---

```{r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
#tidyverse packages
library(here)
library(janitor)
library(dplyr)
library(tidyverse)

#spatial data
library(sf)

#text analysis
library(tidytext)
library(textdata)
library(ggwordcloud)

#print pretty tables
library(gt)
library(kableExtra)

#misc
library(viridis)

```

## Read in data

```{r read in data, message = FALSE, warning = FALSE}
#read in lit review data
data <- read.csv(here("data10.csv")) %>% 
  clean_names() %>% 
  select(-mpa_effect_size) #not currently considering this variable

#read in MEOW TNC Shapefile
meow <- read_sf(here("MEOW-TNC", "meow_ecos.shp")) %>% 
  clean_names()

#make MPA effects definitions dataframe
definitions <- data.frame(mpa_effect = c("Mitigation", "Resilience", "Offset Mortality"),
                          definition = c("MPA stops climate effect directly", "MPA supports more resilient ecosystem", "offsets mortality by restricting industry"))
```

## MPA Effects

Print a table of MPA effects definitions

```{r, message = FALSE, warning = FALSE}
#make table
mpatable <- definitions %>% 
  kbl(col.names = c("MPA Effect",
                    "Definition")
    
  ) %>% 
  kable_styling(full_width = F, html_font = "Arial") %>% 
  row_spec(1:3, color = "black")
  

#print table
mpatable


#save table as .png
save_kable(x = mpatable, file = "MPA Table.png", zoom = 5)

```

Use `ggplot2` to make a bar graph showing the number of occurences of each mpa effect.

```{r mpa effect table, message = FALSE, warning = FALSE}
effect_counts <- data %>% 
  count(mpa_effect)
 
#temp <- effect_counts[-c(1), ]

#create a bar graph to visualize types of mpa effect
mpaeffect_bargraph <- ggplot(data = effect_counts, aes(x = mpa_effect, y = n)) +  #create plot with MPA effect on the x axis and counts on the y-axis
  geom_col(width = 0.75, fill = "#25858EFF") +      #use a bar graph format with slightly narrower bars and fill color steelblue3
  geom_text(aes(label = n), vjust = -.5, size = 5) +    #add labels to columns
  labs(               #rename y-axis, delete x-axis label, and provide a caption
    y = "",
    x = "",
    title = "Number of case studies by type of MPA effect"
    #caption = "Figure 1: Most MPAs either show Effect B (support a more resilient ecosystem) or have no effect."
  ) +
  theme_classic() +
  theme(axis.text = element_text(size = 14, angle = 20, hjust = 1, color = "gray6" ),    #change size, angle, positioning, and color of x-axis tick labels
        plot.caption = element_text(size = 12),        #set font size for caption text
        axis.text.y = element_blank(),      #delete y axis text
        axis.ticks.y = element_blank(),    #delete y axis ticks
        panel.grid.major = element_blank(),    #delete major grid lines
        panel.grid.minor = element_blank(),     #delete minor grid lines
        panel.background = element_rect(fill = "white"),   #make background color white
        axis.line.x = element_blank(),    #delete axis line
        axis.line.y = element_blank(),    #delete axis line
        plot.title = element_text(hjust = 0.5))+ #center title
  expand_limits(y = 29) + #expand limits of y axis so labels on top of the columns don't get cut off
  scale_x_discrete(labels = c("Mitigation", "Resilience", "Offset Mortality", "Negative Effects", "No Effect"))  #rename x-axis labels

mpaeffect_bargraph #prints graph into knitted document

# Save the ggplot as a .png file
ggsave("mpa_effects_bargraph.png", plot = mpaeffect_bargraph, width = 8, height = 4, dpi = 600)

```

Next, I dug into which habitat and geography scenarios were associated with each MPA effect.

```{r, message = FALSE, warning = FALSE}
# negative effect
negative_effect <- data %>% 
  filter(mpa_effect == "negative_effect")

# no effect
no_effect <- data %>% 
  filter(mpa_effect == "no_effect")

# effect_a, mpa stops climate effect directly
effect_a <- data %>% 
  filter(mpa_effect == "effect_a")

# effect b, mpa supports more resilient ecosystem
effect_b <- data %>% 
  filter(mpa_effect == "effect_b")

# effect c, offsets mortality by restricting industry
effect_c <- data %>% 
  filter(mpa_effect == "effect_c")


```

After that, I parsed out which climate effects and geographies most applied to each mpa effect

```{r, message = FALSE, warning = FALSE}
#save list of mpa effects
effect_list <- unique(data$mpa_effect)

# create empty dataframe
mpa_habitats <- data.frame(matrix(ncol = 3, nrow = length(effect_list)))

#provide column names
colnames(mpa_habitats) <- c("mpa_effect", "habitat", "realm")

for(i in 1:length(effect_list)) {  #for every element starting with the first element
  
  #filter by mpa effect and save to holding vector `effect`
  effect <- data %>% filter(mpa_effect == effect_list[i]) 
  
  #make list of all habitats and save to holding vector `habs`
  habs <- unique(effect$habitat_type)   
  
  #delete nas, save to holding vector `habs_no_na`
  habs_no_na <- toString(as.list(habs[!habs == "na"])) #save unique values as a list, and then put the entire list in one cell of the holding vector
  
  #make a list of all marine realms holding vector `re`
  re <- toString(as.list(unique(effect$marine_ecoregion))) #save unique values as a list, and then put the entire list in one cell of the holding vector
  
  #create a holding vector to append as a row to the dataframe
  vec <- c((effect_list[i]),(habs_no_na), (re)) # create vector of mpa effect, habitat, and marine ecoregion
  
  #append rows to dataframe
  mpa_habitats[i,] <- vec
}


#drop rows where mpa_effect cell is blank (delete this line later)
mpa_eff_habitats <- mpa_habitats %>% 
  filter(!(mpa_effect == ""))

#make table
effects_table <- mpa_eff_habitats %>% 
  kbl() %>% 
  kable_styling("hover", full_width = F)

#print table
effects_table


```


## Geographic distribution of existing studies

some code from: https://www.imachordata.com/2015/03/28/meow-its-marine-ecoregions-in-r-2/ 

```{r geography, message = FALSE, warning = FALSE}
location_counts <- data %>% 
  count(marine_ecoregion)

#temp2 <- location_counts[-c(1), ]

#bar plot of country locations
ggplot(data = location_counts, aes(x = marine_ecoregion, y = n)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=.4))

#make spatial polygon for marine realms
realms <- meow %>%
  group_by(realm) %>%
  summarize(geometry = st_combine(geometry)) %>% 
  mutate(realm = recode(realm,
                        "Arctic" = "arctic",
                        "Central Indo-Pacific" = "central_indo_pacific",
                        "Eastern Indo-Pacific" = "eastern_indo_pacific",
                        "Southern Ocean" = "southern_ocean",
                        "Temperate Australasia" = "temperate_autralasia",
                        "Temperate Northern Atlantic" = "temperate_northern_atlantic",
                        "Temperate Northern Pacific" = "temperate_northern_pacific",
                        "Temperate South America" = "temperate_south_america",
                        "Temperate Southern Africa" = "temperate_southern_africa",
                        "Tropical Atlantic" = "tropical_atlantic",
                        "Tropical Eastern Pacific" = "tropical_eastern_pacific",
                        "Western Indo-Pacific" = "western_indo_pacific")) %>% 
  rename(marine_ecoregion = realm)

#create manual coordinates for temperate_northern_pacific centroid
new_coordinates <- st_point(c(-125, 38))

#find centroids of each ecoregion
symbol_pos <- st_centroid(realms, of_largest_polygon = TRUE) %>% 
  assign_in(list(2,7), new_coordinates) #manually replace temperate_northern_pacific centroid

#df <- symbol_pos %>% 
 # filter(marine_ecoregion == "temperate_northern_pacific") %>% 
  #mutate(geometry = new_coordinates)

#join centroids of each ecoregion to counts data, so that the counts data can be placed at the centroid on the map
count_n <- left_join(symbol_pos, location_counts, by = "marine_ecoregion")

#change NAs to 0s
count_n$n[is.na(count_n$n)] <- 0

#make a plot
#refining steps: need to add labels to the bubbles, need to figure out ideal color for bubbles
marine_realms <- ggplot() +
  geom_sf(data = realms, aes(fill = marine_ecoregion, color = NA)) +
  scale_fill_manual(values = c("azure3", "azure3", "azure3", "azure3", "azure3", "azure3", "azure3", "azure3", "azure3", "azure3", "azure3", "azure3")) +
  #scale_fill_grey(start = 0.2, end = 0.8, na.value = "black", aesthetics = "fill" ) +
  #scale_fill_manual(values = viridis::viridis(length(unique(realms$marine_ecoregion)))) + #specifies colors from virdis package manually. this is a workaround to avoid getting the error " Discrete value supplied to continuous scale" when using `scale_fill_viridis`. `scale_fill_gray` worked because its more flexible
  #scale_color_manual(values = viridis::viridis(length(unique(realms$marine_ecoregion)))) +
  labs(
    fill = "",
    caption = "This map does not include 4 global studies and 5 studies in the northern hemisphere."
  ) +
  geom_sf(data = count_n, 
          pch = 21, 
          aes(size = n,
              color = "red"),
          fill = "red",
          show.legend = FALSE) +
  scale_size(range = c(0, 9)) +      #adjust size of bubbles on map
  geom_text(data = count_n, #add a geom for labels of the bubble sizes
            aes(x = st_coordinates(geometry)[,1], y = st_coordinates(geometry)[, 2],   #add x and y values to geom_text
              nudge_x = 0.5, nudge_y = 0.5, label = n)) +    #add data labels to each bubble
  labs(
    x = "",
    y = ""
  ) +
  theme_classic() +
  guides(fill = guide_legend(key_glyph = "point", override.aes = list(color = "white")),  #set legend key as a point, and then change the legend key color to white
         color = FALSE) + #don't show color legend
  #ggtitle("Number of case studies conducted in each marine realm") +
  theme(plot.title = element_text(hjust = 0.5))   #move title to center

marine_realms

ggsave("marine_realms_map.png", plot = marine_realms, width = 8, height = 6, dpi = 800, bg = "transparent")

```

## Type of habitats studied

```{r habitats, message = FALSE, warning = FALSE}
#make a word cloud
 
#make a dataframe with just habitat_types
habitats <- data %>% 
  select(habitat_type) 

#make a dataframe with unnested tokens (one word per row)
habitat_tokens <- habitats %>% 
  unnest_tokens(word, habitat_type) %>% 
  anti_join(stop_words) %>% #remove stopwords
  filter(!(word %in% c("na", "reef", "forest"))) #delete na values and the word reef and forest
         
#count number of every word
habitat_counts <- habitat_tokens %>% 
  count(word) %>% 
  arrange(-n) #order words from most used to least used


```

## Ecosystem characteristics 

## Mechanisms for MPA effectiveness

Make a word cloud of mechanims details for preliminary data vis

```{r mechanisms, message = FALSE, warning = FALSE}
### make a word cloud ###

#make a dataframe with just mechanisms, excluding the rest of the data
mechanisms <- data %>% 
  select(mechanism_details)

#make a dataframe with unnested tokens (one word per row)
mech_tokens <- mechanisms %>% 
  unnest_tokens(word, mechanism_details) %>% #### note to self: correct spelling after uploading new datasheet ####
  anti_join(stop_words) #remove stopwords

#count number of every word
mech_counts <- mech_tokens %>% 
  count(word) %>% 
  arrange(-n) #order words from most used to least used

#make a dataframe of the top 300 words
top_121 <- mech_counts %>% 
  arrange(-n) %>%  #arrange in descending order by count
  slice(1:121)  # keep only the top 121 words (omits all words that only occur once)
  
# make a word cloud
ggplot(data = top_121, aes(label = word)) +
  geom_text_wordcloud(aes(color = n, size = n)) +
  scale_size_area(max_size = 9) + #increases text size
  scale_color_gradientn(colors = c("deepskyblue", "darkorchid3", "turquoise", "palevioletred")) +
  labs(
    title = "Top 121 words describing MPA mechanisms",
    caption = "Omits all words that only occurred once"
  ) +
  theme_minimal()


```
```{r mechanisms, message = FALSE, warning = FALSE}
m_key <- data %>% 
  select(mechanism_key)

key_counts <- m_key %>% 
  count(mechanism_key_action_leads_to_consequence)

```



## Next Steps

- which ecosystems are in each ecoregion?
- which MPA effects are in each ecoregion?
- The distinction between mitigation and resilience is confusing





