---
title: "Chapter 1 Paper Figures - v1"
author: "Mukta Kelkar"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

#tidyverse packages
library(here)
library(janitor)
library(tidyverse)

#spatial data
library(sf)

#datavis
library(RColorBrewer)
library(gridExtra)
library(patchwork)
```


```{r read in data and set up fonts, message=FALSE}
#read in lit review data
data <- read.csv(here("data.csv"), na.strings = c("", "NA")) %>% #read in data with empty cells as NA
  clean_names() %>%
  select(-magnitude) %>% 
  select(-mechanism) %>% 
  mutate(habitat_type = ifelse(habitat_type == "subtropical", "tropical", habitat_type))
  #na.omit() #omit NAs

data_unique <- data %>% 
  select(citation, topic, habitat_type, habitat_subtype, geographic_region) %>% #select variable that are constant within each case study
  distinct(citation, habitat_type, geographic_region, .keep_all = TRUE)

#read in MEOW TNC Shapefile
meow <- read_sf(here("MEOW-TNC", "meow_ecos.shp")) %>% 
  clean_names()

#add google fonts to working memory for figures
#font_add_google(name = "Work Sans",   # Name of the font on the Google Fonts site
#                family = "worksans")

#standard theme for all graphs
my_theme <- theme_classic() +
  theme(base_size = 14)
        #axis.text.x = element_text(angle = 45, hjust = 1))

#source for standard color palettes: https://www.simplifiedsciencepublishing.com/resources/best-color-palettes-for-scientific-figures-and-data-visualizations
#for mpa effect, use six color combination from "Divergent"
mpa_colors <- c("positive_ecosystem" = "#2066a8", 
                "positive_population" = "#cde1ec", 
                "no_effect" = "grey60", 
                "negative_population" = "#f6d6c2", 
                "negative_ecosystem" = "#ae282c")

#for climate effect, use six color combination from "qualitative"
climate_colors <- c("body_size" = "#cecece", 
                    "disease_parasite" = "#a559aa", 
                    "habitat_formation" = "#59a89c", 
                    "mortality" = "#f0c571", 
                    "movement_range" = "#e02b35", 
                    "phenology" = "#082a54", 
                    "reproduction" = "darkolivegreen")

```

## Background 
### Habitat Information
Number of case studies in each habitat type and habitat subtype. Filtered so that each case study is only represented once.
```{r habitat and subhabitat type, message = FALSE}
#figure 1a
habitat_counts <- data %>% 
  count(habitat_type) #make dataset with counts data

tropic_total <- habitat_counts[habitat_counts$habitat_type == "tropical", "n"]
temperate_total <- habitat_counts[habitat_counts$habitat_type == "temperate", "n"]
habitat_total <- habitat_counts %>% summarize(total = sum(n))
tropic_percent <- tropic_total/habitat_total
temperate_percent <- temperate_total/habitat_total

#figure 1a
fig1a <- ggplot(data = habitat_counts, aes (x = reorder(habitat_type, n), y = n)) +
  geom_col(width = 0.7) +
  labs(
    x = "Habitat Type",
    y = ""
  ) +
  my_theme +
  scale_x_discrete(labels = c("tropical" = "Tropical", 
                              "temperate" = "Temperate", 
                              "na" = "NA"),
                   drop = FALSE) +
  coord_flip() +
  ylim(0,130)

#figure 1b
subhabitat_counts <- data %>% 
  count(habitat_subtype) #make dataset with counts data

subhabitat_total <- subhabitat_counts %>% summarize(total = sum(n))
coral_total <- subhabitat_counts[subhabitat_counts$habitat_subtype == "coral_reef", "n"]
coral_percent <- coral_total/subhabitat_total
rockyreef_total <- subhabitat_counts[subhabitat_counts$habitat_subtype == "rocky_reef", "n"]
rockyreef_percent <- rockyreef_total/subhabitat_total
kelp_total <- subhabitat_counts[subhabitat_counts$habitat_subtype == "kelp_forest", "n"]
kelp_percent <- kelp_total/subhabitat_total


#figure 1b
fig1b <- ggplot(data = subhabitat_counts, aes (x = reorder(habitat_subtype, n), y = n)) +
  geom_col(width = 0.7) +
  labs(
    x = "Habitat Subtype",
    y = ""
  ) +
  my_theme +
  scale_x_discrete(labels = c("coral_reef" = "Coral Reef",
                              "rocky_reef" = "Rocky Reef",
                              "na" = "NA",
                              "kelp_forest" = "Kelp Forest",
                              "shallow_reef" = "Shallow Reef",
                              "seagrass" = "Seagrass",
                              "deep_reef" = "Deep Reef",
                              "intertidal" = "Intertidal"),
                   drop = FALSE) +
  coord_flip() +
  ylim(0,119)



fig1 <- wrap_plots(fig1a, fig1b)

#print figure
fig1
```

### Assemblage Information
How much each assemblage type was studied proportionally. I don't know if it would make sense to manually group some of the assemblage types for data analysis/more refined visuals (I did this for the metrics later on).
```{r assemblages and species, fig.width=6, fig.height=9}
assemblage_all <- data %>% 
  count(assemblage_type) %>% 
  mutate(proportion = n /sum(n))

unique_assemblages <- data %>% 
  distinct(assemblage_type)

#write_csv(unique_assemblages, "unique_assemblages.csv")

lookup_table_assembl <- read.csv(here("unique_assemblages.csv"))

assemblages_combined <- assemblage_all %>% 
  left_join(lookup_table_assembl, by = "assemblage_type") %>% 
  select(n, grouped_assemblage_highest) %>% 
  group_by(grouped_assemblage_highest) %>% 
  summarise(total = sum(n), .groups = "drop") %>% 
  ungroup() %>% 
  mutate(proportion = (total / sum(total)))

#figure - first option
fig_assemblage <- ggplot(data = assemblages_combined, aes (x = reorder(grouped_assemblage_highest, total), y = total)) +
  geom_col() +   #set up bar chart
  labs(
    x = "Assemblage Type",
    y = ""
  ) +
  my_theme +
  scale_x_discrete(labels = c("fish" = "Fish",
                              "na" = "NA",
                              "coral" = "Coral",
                              "benthic" = "Benthic",
                              "invertebrates" = "Invertebrates",
                              "macroalgae" = "Macroalgae"),
                   drop = FALSE) +
  coord_flip() +
  ylim(0,119)

  

figure_1 <- (fig1a / fig1b / fig_assemblage) +
  theme(axis.title.x = element_blank()) +
  plot_annotation(caption = "Number of Occurrances")

figure_1

```

```{r effect on target fish}
assemblages_target <- data %>% 
  select(assemblage_type, climate_effect, mpa_effect)

lookup_table_assembl <- read.csv(here("unique_assemblages.csv"))

assemblages_target_simple <- left_join(assemblages_target, lookup_table_assembl, by = "assemblage_type") %>% 
  select(!c(grouped_assemblage_highest, grouped_assemblage_medium, assemblage_type))

target_effect_counts <- assemblages_target_simple %>% 
  group_by(mpa_effect, target_fish) %>% 
  count(target_fish) %>% 
  ungroup() %>% 
  mutate(proportion = n /sum(n)) %>% 
  filter(target_fish %in% c("target_fish", "non-target_fish"))
```


### Geography
Bar plot and map of which marine realms (defined by Marin Ecoregions of the World) the case studies were in. Filtered so that each case study is only represented once.
```{r figure 2: map of study sites, message = FALSE}
#ecoregion_counts <- data_unique %>% 
#  count(geographic_region) #make dataset with counts data

#ggplot(data = ecoregion_counts, aes(x = reorder(geographic_region, -n), y = n)) +
#  geom_col() +
#  theme(axis.text.x = element_text(angle = 30, vjust = 1, hjust=1))

#fig1c_bar <- ggplot(data = ecoregion_counts, aes (x = reorder(geographic_region, -n), y = n)) +
#  geom_bar(stat = "identity", width = .75, color = "white") +   #set up bar chart
#  theme_classic(base_size = global_size) +
#  labs(
#    x = "Ecoregion",
#    y = "Number of Occurences"
#  ) +
#  theme(
#    axis.text.x = element_text(angle = 45, hjust = 1)
#  )

#make spatial polygon for marine realms
#realms <- meow %>%
#  group_by(realm) %>%
#  summarize(geometry = st_combine(geometry)) %>% 
#  mutate(realm = recode(realm,
#                        "Arctic" = "arctic",
#                        "Central Indo-Pacific" = "central_indo_pacific",
#                        "Eastern Indo-Pacific" = "eastern_indo_pacific",
#                        "Southern Ocean" = "southern_ocean",
#                        "Temperate Australasia" = "temperate_autralasia",
#                        "Temperate Northern Atlantic" = "temperate_northern_atlantic",
#                        "Temperate Northern Pacific" = "temperate_northern_pacific",
#                        "Temperate South America" = "temperate_south_america",
#                        "Temperate Southern Africa" = "temperate_southern_africa",
#                        "Tropical Atlantic" = "tropical_atlantic",
#                        "Tropical Eastern Pacific" = "tropical_eastern_pacific",
#                        "Western Indo-Pacific" = "western_indo_pacific")) %>% 
#  rename(marine_ecoregion = realm)

#create manual coordinates for temperate_northern_pacific centroid
#new_coordinates <- st_point(c(-125, 38))

#find centroids of each ecoregion
#symbol_pos <- st_centroid(realms, of_largest_polygon = TRUE) %>% 
#  assign_in(list(2,7), new_coordinates) %>% #manually replace temperate_northern_pacific centroid
#  rename(geographic_region = marine_ecoregion)

#df <- symbol_pos %>% 
 # filter(marine_ecoregion == "temperate_northern_pacific") %>% 
  #mutate(geometry = new_coordinates)

#join centroids of each ecoregion to counts data, so that the counts data can be placed at the centroid on the map
#count_n <- left_join(symbol_pos, ecoregion_counts, by = "geographic_region")

#change NAs to 0s
#count_n$n[is.na(count_n$n)] <- 0

#make a plot
#refining steps: need to add labels to the bubbles, need to figure out ideal color for bubbles
#marine_realms <- ggplot() +
#  geom_sf(data = realms, aes(fill = marine_ecoregion, color = NA)) +
#  scale_fill_manual(values = viridis::viridis(length(unique(realms$marine_ecoregion)))) + #specifies colors from virdis package manually. this is a workaround to avoid getting the error " Discrete value supplied to continuous scale" when using `scale_fill_viridis`. `scale_fill_gray` worked because its more flexible
  #scale_color_manual(values = viridis::viridis(length(unique(realms$marine_ecoregion)))) +
#  labs(
#    fill = "",
#    caption = "This map does not include 4 global studies and 5 studies in the northern hemisphere."
#  ) +
#  geom_sf(data = count_n, 
#          pch = 21, 
#          aes(size = n,
#              color = "red"),
#          fill = "red",
#          show.legend = FALSE) +
#  scale_size(range = c(0, 9)) +      #adjust size of bubbles on map
#  geom_text(data = count_n, #add a geom for labels of the bubble sizes
#            aes(x = st_coordinates(geometry)[,1], y = st_coordinates(geometry)[, 2],   #add x and y values to geom_text
#              nudge_x = 0.5, nudge_y = 0.5, label = n)) +    #add data labels to each bubble
#   labs(
#     x = "",
#     y = ""
#   ) +
#   theme_classic() +
#   guides(fill = guide_legend(key_glyph = "point", override.aes = list(color = "white")),  #set legend key as a point, and then change the legend key color to white
#          color = FALSE) + #don't show color legend
#   ggtitle("Number of case studies conducted in each marine realm") +
#   theme(plot.title = element_text(hjust = 0.5))   #move title to center
# 
# marine_realms
```

## Map
```{r figure 2, fig.width=9, fig.height=5}
map_latlong <- data %>% 
  select(climate_effect, mpa_effect, latitude, longitude) %>% 
  mutate(
    jitter_lat = jitter(latitude, amount = 5),
    jitter_long = jitter(longitude, amount = 5)
  )

#count number of occurrences in western US coast
west_US_bbox <- map_latlong %>% 
  filter(latitude < 48.5, latitude > 32.5, 
         longitude < -117, longitude > -126) %>% 
  mutate(count = n())

#count number of occurrences in Caribbean coast
carribean_bbox <- map_latlong %>% 
  filter(latitude < 27, latitude > 7, 
         longitude < -58.7, longitude > -91) %>% 
  mutate(count = n())
  
#count number of occurrences in European coast
europe_bbox <- map_latlong %>% 
  filter(latitude < 68, latitude > 26.8, 
         longitude > -28.7, longitude < 58.1) %>% 
  mutate(count = n())

#count number of occurrences in SE Asia coast
se_asia_bbox <- map_latlong %>% 
  filter(latitude < 28.7, latitude > -11.2, 
         longitude > 86.4, longitude < 160) %>% 
  mutate(count = n())

#count number of occurrences in Australian coast
australia_bbox <- map_latlong %>% 
  filter(latitude < -10.3, latitude > -38.7, 
         longitude > 109.1, longitude < 160) %>% 
  mutate(count = n())

#count number of occurrences in Tasmanian coast
tasmania_bbox <- map_latlong %>% 
  filter(latitude < -40, latitude > -44, 
         longitude > 143.8, longitude < 149.2) %>% 
  mutate(count = n())

# Get world map data
world_map <- map_data("world")

map_climate <- ggplot() +
  geom_polygon(data = world_map, aes(x = long, y = lat, group = group),
               fill = "gray90", color = NA) +
  geom_point(data = map_latlong, aes(x = jitter_long, y = jitter_lat, color = climate_effect), size = 2) +
  coord_fixed(1.3) +  # Keeps aspect ratio of map
  theme_minimal() +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text = element_blank()) +
  labs(title = "Geographic Distribution of Climate Effect",
       color = "Climate Effect") +
  scale_color_discrete(labels = c("body_size" = "Body Size",
                                 "disease_parasite" = "Disease & Parasitism",
                                 "habitat_formation" = "Habitat Formation",
                                 "mortality" = "Mortality",
                                 "movement_range" = "Movement & Range Shifts",
                                 "phenology" = "Phenology",
                                 "reproduction" = "Reproduction")) +
  scale_color_manual(values = climate_colors)

map_mpa <- ggplot() +
  geom_polygon(data = world_map, aes(x = long, y = lat, group = group),
               fill = "gray90", color = NA) +
  geom_point(data = map_latlong, aes(x = jitter_long, y = jitter_lat, color = mpa_effect), size = 2) +
  coord_fixed(1.3) +  # Keeps aspect ratio of map
  theme_minimal() +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  labs(title = "Geographic Distribution of MPA Effect",
       color = "MPA Response Type",
       caption = "Figure ") +
  scale_color_manual(values = mpa_colors)


map_climate
map_mpa

#ggplot() +
#  geom_polygon(data = world_map, aes(x = long, y = lat, group = group),
#               fill = "gray90", color = NA) +
#  geom_point(data = map_latlong, aes(x = longitude, y = latitude, color = climate_effect), size = 2) +
#  scale_color_viridis_d(option = "magma") +
#  coord_fixed(1.3) +  # Keeps aspect ratio of map
#  theme_minimal() +
#  theme(axis.title.x = element_blank(),
#        axis.title.y = element_blank(),
#        axis.text = element_blank()) +
#  labs(title = "Geographic Distribution of Climate Effect",
#       color = "Climate Effect")

#ggplot() +
#  geom_polygon(data = world_map, aes(x = long, y = lat, group = group),
#               fill = "gray90", color = NA) +
#  geom_jitter(data = map_latlong, aes(x = longitude, y = latitude, color = climate_effect), 
#              width = 5, height = 5, size = 1.75) +
#  scale_color_viridis_d(option = "magma") +
#  coord_fixed(1.3) +  # Keeps aspect ratio of map
#  theme_minimal() +
#  theme(axis.title.x = element_blank(),
#        axis.title.y = element_blank(),
#        axis.text = element_blank()) +
#  labs(title = "Geographic Distribution of Climate Effect",
#       color = "Climate Effect")
```

## Frameworks
### MPA  & Climate Effects
Proportion of occurrences of each MPA effect.
```{r mpa effects, message = FALSE}
#figure 3a
mpa_effect_counts <- data %>% 
  count(mpa_effect) %>% #make dataset with counts data
  mutate(proportion = n /sum(n))

#figure 3a
fig3a <- ggplot(data = mpa_effect_counts, aes (x = reorder(mpa_effect, n), y = n, fill = mpa_effect)) +
  geom_bar(stat = "identity", width = .75) +   #set up bar chart
  labs(
    x = "MPA Response Type",
    y = "Number of Occurences"
  ) +
  scale_x_discrete(labels = c("no_effect" = "No Effect",
                              "positive_ecosystem" = "Ecosystem (+)", 
                              "negative_ecosystem" = "Ecosystem (-)",
                              "positive_population" = "Population (+)", 
                              "negative_population" = "Population (-)",
                              "not_assessed" = "Not Assessed")) +
  scale_fill_manual(values = mpa_colors) +
  my_theme +
  coord_flip()

fig3a
```

Proportion of occurrences of each climate effect.
```{r climate effects, message = FALSE, fig.width=8, fig.height=6}
#figure 3b data
climate_counts_prelim <- data %>% 
  count(climate_effect) %>% #preliminary counts of climate effect (without separating rows with multiple categories)
  mutate(proportion = n /sum(n))

#figure 3b ggplot
fig3b <- ggplot(data = climate_counts_prelim, aes (x = reorder(climate_effect, n), y = n, fill = climate_effect)) +
  geom_bar(stat = "identity", width = .75) +   #set up bar chart
  labs(
    x = "Ecological Impact of Climate Change",
    y = "Number of Occurences"
  ) +
  my_theme +
  scale_fill_manual(values = climate_colors) +
  coord_flip()

fig3b #view figure

figure_3 <- (fig3a / fig3b) +
  theme(axis.title.x = element_blank()) +
  plot_annotation(caption = "Figure #: Number of occurrances of each MPA response type and each climate impact.")

figure_3


```

### Heatmap 
```{r figure 5: heatmap, message=FALSE}
eco_effects <- data %>% 
  select(mpa_effect, climate_effect) %>% 
  group_by(mpa_effect, climate_effect) %>% 
  count(mpa_effect, climate_effect, name = "count")

heatmap_v2 <- ggplot(data = eco_effects, aes(x = climate_effect, y = mpa_effect, fill = count)) +
  geom_tile() +
  labs(
    x = "Climate Impact",
    y = "MPA Response",
    caption = str_wrap("Figure #: Number of occurrences for each combination of climate impact and MPA response. Color darkness corresponds to frequency with which each combination was observed.", width = 70)
  ) +
  #scale_y_discrete(labels = c("Ecosystem (+)", "Population", "Ecosystem (-)", "None")) +
  scale_fill_gradient2(low = "grey",
                      high = "black") +
  theme_classic(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.caption = element_text(hjust = 0, margin = margin(t = 5)),
  )

heatmap_v2

ggsave("heatmap_v2.png", plot = heatmap_v2, width = 10, height = 5, bg = "transparent")
```

### MPA Effects by habitat type
```{r}
mpa_effects_by_habitat <- data %>% 
  select(habitat_type, habitat_subtype, climate_effect, mpa_effect) %>% 
  filter(!mpa_effect == "not_assessed") %>% 
  filter(!habitat_subtype == "na") %>% 
  filter(!habitat_type == "na") %>% 
  group_by(habitat_type, habitat_subtype) %>% 
  count(mpa_effect) %>% 
  mutate(proportion = n /sum(n))

temperate <- mpa_effects_by_habitat %>% 
  filter(habitat_type == "temperate")

fig4 <- ggplot(data = mpa_effects_by_habitat, aes(fill = mpa_effect, y = n, x = habitat_subtype)) + 
  geom_bar(position = 'stack', stat = 'identity') + 
  facet_wrap(vars(habitat_type)) +
  theme_classic() +
  labs(
    fill = "MPA Effect"
  ) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  scale_fill_manual(values = mpa_colors)

ggplot(data = mpa_effects_by_habitat, aes(fill = mpa_effect, y = proportion, x = habitat_subtype)) + 
  geom_bar(position = 'stack', stat = 'identity') + 
  facet_wrap(vars(habitat_type)) +
  theme_classic() +
  scale_fill_manual(values = mpa_colors)

fig4
```



### Metrics 
Metrics used to measure each MPA effect type, not normalized (I'm planning to, I just ran out of time today.)
```{r figure 5: metrics to measure, message=FALSE, fig.width=9, fig.height=10}
metrics_measured <- data %>% 
  select(climate_effect, ecological_fitness_metric) %>% 
  group_by(climate_effect, ecological_fitness_metric) %>% 
  count(climate_effect, ecological_fitness_metric, name = "count")

#unique_metrics <- data %>% 
  #distinct(ecological_fitness_metric)

#write_csv(unique_metrics, "unique_metrics.csv")

lookup_table <- read.csv(here("unique_metrics.csv"))

metrics_combined <- metrics_measured %>% 
  left_join(lookup_table, by = "ecological_fitness_metric") %>% 
  group_by(climate_effect, grouped_metric) %>% 
  summarise(total_count = sum(count), .groups = "drop")
  
fig5 <- ggplot(data = metrics_combined, aes(x = grouped_metric, y = total_count, fill = )) +
  geom_bar(stat = "identity") +
  facet_wrap(~climate_effect, scales = "free_x") +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(
    x = "Ecological Fitness Metric",
    y = "Number of Occurences"
  )

fig5

```


## Print and Save Figures
```{r print figures, eval = FALSE}
#print figures
figure_1
map_climate
map_mpa
figure_3
heatmap_v2
fig4
fig5

```


```{r}
data_neg <- data %>% 
  filter(mpa_effect %in% c("negative_population", "negative_ecosystem"))

data_pos <- data %>% 
  filter(mpa_effect %in% c("positive_population", "positive_ecosystem"))

data_null <- data %>% 
  filter(mpa_effect == "no_effect")
```


```{r save figures, eval=FALSE}
# #save figures
# ggsave("fig1_combined.png", plot = fig1, width = 10, height =5, dpi = 300)
# ggsave("fig1a_bar.png", plot = fig1a_bar, width = 9, height = 5, bg = "transparent")
# ggsave("fig1a_sub_bar.png", plot = fig1a_sub_bar, width = 9, height = 5, bg = "transparent")
# ggsave("fig1b_bar.png", plot = fig1b_bar, width = 9, height = 5, bg = "transparent") #save fig1b with a transparent background
# ggsave("fig1c_bar.png", plot = fig1c_bar, width = 9, height = 5, bg = "transparent")
# ggsave("fig1d_bar.png", plot = fig1d_bar, width = 9, height = 5, bg = "transparent")
# #ggsave("fig1e_bar.png", plot = fig1e_bar, width = 9, height = 5, bg = "transparent")
```




